---
title: "20382424 coursework"
output: html_document
---

This is a R markdown file for the statistical analysis oj Vunder drug trial.
First we read the datasets provided to use using read.csv().
```{r}
data <- read.csv("Cwk_data_20382424.csv") # Import data
rectime<-read.csv("Cwk_rectime_20382424.csv") #Import data
```

**Task 1**
Let µ denote the mean effect of Vunder. Calculate the sample size required to reject the null hypothesis H0 : µ = 122 in favour of the alternative hypothesis H1 : µ = 113 with significance level 5% and power 85%.
```{r}
mu<-9 # Calculate mu by subtracting alternative hypothesis from the null hypothesis.
sigma<-24 # 24 hours is the standard deviation for recovery time of Xanadu.
alpha<-0.05 #significance level of 5%.
power<-0.85 #power is given as 85%.
sample_size<-((qnorm(1-alpha)-qnorm(1-power))*sigma/mu)^2 # calculate the sample size 

```

**Task 2**
Construct a 99% confidence interval for µ. Describe how the confidence interval is constructed.
In this task, we use the 'rectime' data which consist of the recovery time of patients in hospital Atesta drug trial.
```{r}
n=80 #80 is the given sample size.

t.test(rectime,mu=122, conf.level = 0.99) # t-test to find the 99% confidence interval.

```
**Task 3**
Create a covariate bmi. The bmi of a patient is given by (weight in kg)/(height in m)^2.
There are four main categories of bmi underweight (< 18.5), healthy (18.5 − 25), overweight (25 − 30) and obese (> 30). Is the bmi category of a patient receiving Vunder independent of hospital?
```{r}

data$bmi<-(data$weight/(data$height/100)^2) #calculate BMI and store it into a variable
hist(data$bmi,xlab="BMI",main="Histogram for BMI",freq=FALSE) # create histogram for BMI.


# BMI is continuous variable,we convert it the values to discrete by forming groups.
#Map underweight,healthy,overweight,obese as 0,1,2,3 respectively and store it in a new variable. 

data$bmicategory=cut((data$weight/(data$height/100)^2), br=c(0,18.5,25,30,Inf),labels=c(0,1,2,3)) 

test<-table(data$bmicategory,data$hospital) #create contingency table
test #print the table

qchisq(1-0.05,df=9)#calculate chi-square distribution.
chisq.test(test) #performed chisq test
```

**Task 4**
Find the most appropriate linear regression model for predicting the recovery time after
surgery, y, in terms of the covariates provided, including bmi.
Hint: Check that categorical variables are correctly coded in your data.

```{r}

data$bmicategory=as.numeric(levels(data$bmicategory))[data$bmicategory] #convert 'bmicategory' covariate into numeric.

model=lm(Recovery.time~.,data = data) #create a model
summary(model) #summary statistics of the model

```
```{r}
bestmodel=step(model) #step function for calculating the AIC values

summary(bestmodel)#summary statistics for bestmodel
```
Use your model to predict the recovery time of a 41 year old female who does not smoke in hospital 4 given that their height is 155 cm and their weight is 67.7 kg.

```{r}
age=41
sex=0
weight=67.7
height=155
smoker=0
hospital=4
bmi=weight/(height)^2
test_data=data.frame(age,sex,weight,height,smoker,hospital,bmi)
y_pred=predict(bestmodel,test_data)
y_pred
```

```{r}
#Histogram for residuals to check for normal distribution
hist(bestmodel$residuals,xlab="Residuals",main="Histogram for residuals",freq=FALSE)
curve(dnorm(x,0,sd(bestmodel$residuals)),-30,30,col=2,add=TRUE) 
```

```{r}
#Scatter plot for fitted vales vs Residual
bestmodel$fitted.values
plot(bestmodel$fitted.values,bestmodel$residuals,xlab="Fitted values",ylab="Residuals", main="Plot for Fitted values vs Residuals")
abline(h=0,lty=2)
```